---
title: "hw6"
output: github_document
---


```{r setup, include=FALSE}
library(tidyverse)
library(modelr)
library(mgcv)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	message = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```
### Problem 1:
To obtain a distribution for $\hat{r}^2$, we'll follow basically the same procedure we used for regression coefficients: draw bootstrap samples; the a model to each; extract the value I'm concerned with; and summarize. Here, we'll use `modelr::bootstrap` to draw the samples and `broom::glance` to produce `r.squared` values. 

```{r weather_df, cache = TRUE}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```


```{r}
weather_df %>% 
  modelr::bootstrap(n = 1000) %>% 
  mutate(
    models = map(strap, ~lm(tmax ~ tmin, data = .x) ),
    results = map(models, broom::glance)) %>% 
  select(-strap, -models) %>% 
  unnest(results) %>% 
  ggplot(aes(x = r.squared)) + geom_density()
```

In this example, the $\hat{r}^2$ value is high, and the upper bound at 1 may be a cause for the generally skewed shape of the distribution. If we wanted to construct a confidence interval for $R^2$, we could take the 2.5% and 97.5% quantiles of the estimates across bootstrap samples. However, because the shape isn't symmetric, using the mean +/- 1.96 times the standard error probably wouldn't work well.

We can produce a distribution for $\log(\beta_0 * \beta1)$ using a similar approach, with a bit more wrangling before we make our plot.

```{r}
weather_df %>% 
  modelr::bootstrap(n = 1000) %>% 
  mutate(
    models = map(strap, ~lm(tmax ~ tmin, data = .x) ),
    results = map(models, broom::tidy)) %>% 
  select(-strap, -models) %>% 
  unnest(results) %>% 
  select(id = `.id`, term, estimate) %>% 
  pivot_wider(
    names_from = term, 
    values_from = estimate) %>% 
  rename(beta0 = `(Intercept)`, beta1 = tmin) %>% 
  mutate(log_b0b1 = log(beta0 * beta1)) %>% 
  ggplot(aes(x = log_b0b1)) + geom_density()
```

As with $r^2$, this distribution is somewhat skewed and has some outliers. 

The point of this is not to say you should always use the bootstrap -- it's possible to establish "large sample" distributions for strange parameters / values / summaries in a lot of cases, and those are great to have. But it is helpful to know that there's a way to do inference even in tough cases. 

### Problem 2:
Importing the raw homicides dataset:
```{r, message=FALSE}
library (readr)

urlfile="https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv"

homicide_raw <-read_csv(url(urlfile)) %>% 
  janitor::clean_names()
```
```{r }
homicide = homicide_raw %>%
unite('city_state',"city":"state", remove = FALSE)%>%
  select(-city,-state)%>%
   mutate (resolved = case_when (disposition == "Closed without arrest" ~ "unresolved", 
                                  disposition == "Open/No arrest" ~ "unresolved", 
                                  disposition == "Closed by arrest" ~ "resolved"),
           resolved = as.factor(resolved),
           resolved = fct_relevel(resolved, "unresolved"),
           victim_age = as.numeric(victim_age)) %>%
    filter(!(city_state %in% c("Dallas_TX" , "Phoenix_AZ","Kansas City_MO",  "Tulsa_AL"))) %>% 
     filter(victim_race == "White" | victim_race == "Black")
```

```{r}

baltimore = homicide %>%
  filter(city_state == "Baltimore_MD")

balt_OR = baltimore %>% 
  glm(resolved ~ victim_age+victim_race+victim_sex, data = ., family=binomial()) %>% 
  broom::tidy() %>% 
  mutate(OR = exp(estimate),
         L_95CI = exp(estimate - (1.96*std.error)),
         U_95CI = exp(estimate +(1.96*std.error))) %>% 
  filter(term== "victim_sexMale") %>% 
  select(OR, L_95CI, U_95CI)

knitr::kable(balt_OR)
```

The estimate of the adjusted odds ratio for solving homicides comparing male victims to female victims keeping all other variables fixed are 0.43 with a 95% CI (0.32, 0.56)

```{r}
OR_city = homicide%>% 
  nest(data = -city_state) %>% 
  mutate(model_city = map(data, ~ glm(resolved ~ victim_age + victim_sex + victim_race, 
                                 data = ., family = "binomial")), 
         result = map(model_city, broom::tidy)) %>% 
  select(-data, -model_city) %>% 
  unnest(result) %>% 
  mutate(OR = exp(estimate),
         L_95ci = exp(estimate - 1.96*std.error),
         U_95ci = exp(estimate + 1.96*std.error)) %>% 
  filter(term == "victim_sexMale") %>% 
  select(city_state, OR, L_95ci, U_95ci)
```

```{r}
plot_city = OR_city %>% 
  mutate(city_state = fct_reorder(city_state, OR)) %>% 
  ggplot(aes(x = city_state, y = OR, colour = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = L_95ci, ymax = U_95ci, width = .3)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1), 
        legend.position = "none") + 
  labs(
    x = "City",
    y = "Estimated odds ratio",
    title = " Adjusted odds of solving homicides compaing male vs female victim across major U.S. cities, 2007-2017",
    caption = paste0(
          "Error bars represent the 95% confidence interval.",
          "\nSource: The Washington Post."))
plot_city
```


Based on the graph, it can be noted that across most cities the ORs are less than 1. This suggests that there are decreased odds of resolving a homicide when the victim is male vs. when the victim is female. This odds are lowest for New York. For many cities the 95% CI fo the adjusted OR crosses the null value of , which denotes a statistically insignificant association between sex and solving homicides.The only cities which have increased odds of solving a homicide when the victim is male vs. when the victim is female are Albequerque, Stockton and Fresno but their 95% CI crosses the null value of 1.


### Problem 3

```{r}

birthweight = read_csv("./data/birthweight.csv") %>%
  janitor::clean_names() %>% 
  mutate(bwt = bwt/453.59) %>% 
    mutate(
   babysex = recode(babysex, `1` = "male", `2` = "female"),
    babysex = factor(babysex)
)
   
```
For tidying the dataset, I converted birthweight of the baby from grams to pounds to maintain consistency in measuring weight between variables bwt (birthweight of baby) and delwt (mother's weight at delivery). I also converted babysex to a factor variable.

Proposing my own model:
First I included variables associated with the baby that I think are helpful in predicting baby's weight
```{r}
my_model_1 = lm(bwt~ babysex +bhead+blength+gaweeks, data=birthweight)
 summary(my_model_1)

```

After comparing the estimates, it can be observed that the p value of all variables is < .05 and is significant. Moreover the $r^2$ value is 0.6891 indicating that 68.91% of variance in the birthweight can be explained by variability in the selected variables.

I also want to add variables associated with the baby's parents that I think are important in predicting baby's weight

```{r}
my_model_2 = lm(bwt~ babysex +bhead+blength+gaweeks+delwt+smoken+fincome, data=birthweight)
summary(my_model_2)
```

It can be seen in the summary of the 2nd model that the p value of all variables > 0.05 and is significant. Moreover there is an increase the $r^2$ value of the model i.e 0.7013

```{r residual pred plots}
bw_df = birthweight %>% 
  modelr::add_residuals(my_model_2) %>% 
  modelr::add_predictions(my_model_2) 

bw_df %>% 
  ggplot(aes(x = pred, y = resid)) + geom_point(alpha = 0.5, size = 0.8, colour = "blue") + 
  labs(
    title = " residuals vs. predicted values",
    x = "Predicted value",
    y = "Residual")

```
It can be observed that most points are scattered around residuals -2 to +2 with majority being around residual 0 and the model is not violating assumptions. There are outlier residual values associated with smaller predicted values which should be checked.

Based on the summary statistics and the plot I am selecting my_model_2 to compare with the 2 models in Q3 of hw6


Cross validation:
```{r}
my_model = lm(bwt~ babysex +bhead+blength+gaweeks+delwt+smoken+fincome, data=birthweight)
hw_model_1 = lm(bwt ~ blength + gaweeks, data = birthweight)
hw_model_2 = lm(bwt ~ bhead + blength +babysex + bhead*blength*babysex, data =birthweight) 
 
```

```{r}

bw_ds =
  crossv_mc(birthweight, 100) %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))


```

```{r}
bw_ds = 
  bw_ds %>% 
  mutate(
    my_model  = map(train, ~lm(bwt~ babysex +bhead +blength +gaweeks +delwt +smoken +fincome, data = .x)),
    hw_model_1= map(train, ~lm(bwt ~ blength + gaweeks, data = .x)),
    hw_model_2= map(train, ~lm(bwt ~ bhead + blength +babysex + bhead*blength*babysex, data = as_tibble(.x)))) %>% 
  mutate(
    rmse_my_model = map2_dbl(my_model, test, ~rmse(model = .x, data = .y)),
    rmse_hw_model_1 =map2_dbl(hw_model_1, test, ~rmse(model = .x, data = .y)),
    rmse_hw_model_2= map2_dbl(hw_model_2, test, ~rmse(model = .x, data = .y)))
```


```{r}
bw_ds %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin()
```


Based on the graph, my model has the best prediction accuracy with the lowest RMSE values. This is followed by hw_model 2 which included variables bhead, blength, babysex and their interaction terms. The largest RMSE value was found for hw_model_1 which included variables blength and gaweeks.






